---
title: "OMNR_Precipitation_Summary_Stats"
author: "Thomas Van Zuiden"
date: "August 24, 2016"
output: html_document
---
### Summarising Projected Precipitation data across 19 GCMs for 149 lakes

#### 1. Install packages
```{r}
########### PRECIPITATION SUMMARY FOR 149 ONTARIO LAKES ##############

install.packages("dpylr")
library(dplyr)
install.packages("tidyr")
library(tidyr)
install.packages("reshape")
library(reshape)
```


#### Method 1
###### Possible if all data to be explored have common characters in their column headers

```{r}
setwd("C:\\Users\\adminuser\\Desktop\\EVERYTHING THOMAS\\OMNR Alagae project")
x<-read.csv("SC_Ontario+Algonqiun+Algoma_Future_clim.csv")

#Convert data to tbl class
tbl_df(x)
#Select a subset of variables by column
x1 <- select(x, LAKE.NUMBER, contains("prec_"), contains("26pr70"), 
             contains("45pr70"), contains("60pr70"), contains("85pr70"))
names(x1)
#Convert tbl class back to a data frame
x2<-data.frame(x1)
names(x2)
write.csv (x2, file = "algalprecip_boxplot_data.csv",row.names=FALSE)
```

#### Method 2
###### Extremely tedious... only useful for when extracting columns with highly specific names
```{r}
setwd("C:\\Users\\adminuser\\Desktop\\EVERYTHING THOMAS\\OMNR Alagae project")
x<-read.csv("SC_Ontario+Algonqiun+Algoma_Future_clim.csv")
#RCP2.6
x1 <- x[, c(9:56, 249:260, 633:644, 969:980, 1305:1316, 1689:1700, 2073:2084, 2457:2468,
            3033:3044, 3609:3620, 3993:4004, 4329:4340, 4617:4628, 4905:4916, 5337:5348,
            5721:5732)]
#RCP4.5
x2 <- x[, c(297:308, 681:692, 1017:1028, 1353:1364, 1737:1748, 2121:2132, 2505:2516, 2745:2756,
            3081:3092, 3321:3332, 3657:3668, 4041:4052, 4377:4388, 4665:4676, 4953:4964, 5097:5108,
            5385:5396, 5769:5780, 6009:6020)]
#RCP6.0
x3 <- x[, c(345:356, 729:740, 1401:1412, 1785:1796, 2169:2180, 2553:2564, 3129:3140, 3705:3716,
            4089:4100, 4425:4436, 5433:5444, 5817:5828)]
#RCP8.5
x4 <- x[, c(393:404, 777:788, 1065:1076, 1449:1460, 1833:1844, 2217:2228, 2601:2612, 2793:2804,
            3177:3188, 3369:3380, 3753:3764, 4137:4148, 4713:4724, 5001:5012, 5481:5492,
            5865:5876, 6057:6068)]
x5<-data.frame(x1,x2,x3,x4)
write.csv (x5, file = "algalprecip_boxplot_data1.csv",row.names=FALSE)
```

#### Formatting data and calculating some summary stats
```{r}
x<-read.csv("algalprecip_boxplot_data.csv")
tbl_df(x)
names(x)

#I am selecting climate values for winter months only
##NOTE: cannot use "contains" function with prec_1 or any pr701 because I do not want prec_10 or pr7010

x1 <- select(x, contains("prec_12"), ends_with("prec_1"), contains("prec_2"),
             contains("26pr7012"), ends_with("26pr701"), contains("26pr702"),
             contains("45pr7012"), ends_with("45pr701"), contains("45pr702"),
             contains("60pr7012"), ends_with("60pr701"), contains("60pr702"),
             contains("85pr7012"), ends_with("85pr701"), contains("85pr702"))

x2 <- stack(x1)
x2 <- rename(x2, c(values = "Total_Winter_Precipitation",ind = "GCM"))
names(x2)

x2[grep("prec_*", as.character(x2$GCM)), "Climate_Scenario"] <- "1950-2000 Prec"
x2[grep("*26pr70*", as.character(x2$GCM)), "Climate_Scenario"] <- "RCP 2.6"
x2[grep("*45pr70*", as.character(x2$GCM)), "Climate_Scenario"] <- "RCP 4.5"
x2[grep("*60pr70*", as.character(x2$GCM)), "Climate_Scenario"] <- "RCP 6.0"
x2[grep("*85pr70*", as.character(x2$GCM)), "Climate_Scenario"] <- "RCP 8.5"
### Check to see if the new column has labelled the rows appropriately with "View(x2)"

x2 %>%  #Get summary stats on each scenario (mean,  St. Dev.,  Coef. of Variation)
  group_by(Climate_Scenario) %>% 
  summarize(mean.wint.prec=mean(Total_Winter_Precipitation), sd.wint.prec=sd(Total_Winter_Precipitation)) %>%  
  mutate(coef=mean.wint.prec/sd.wint.prec) ## Precipitation increases w. climate change in this region
```

#### Visualisation with box plots
```{r}
boxplot(Total_Winter_Precipitation~Climate_Scenario,data=x2, 
        xlab="Climate Change Scenario", ylab="Total Winter Precipitation (mm)",
        cex.axis = 1.2, cex.lab = 1.2)
```

#### Apply same procedure to other seasons, and for temperature values as well